---
title: "Tofactinib UC"
author: "Andrew Y.F. Li Yim"
date: "2026-05-27"
output: html_document
---

```{r library, include = FALSE}
library(dplyr)
library(DESeq2)
library(ggplot2)
library(ggrastr)
library(readxl)
library(readr)
```

```{r paths, include = FALSE}
degs_TOF_BINF_R_w8vw0_own_csv <- "output/260519/rna_Tissue_R_Inflamed_Week_8_vs_Baseline_deseq2_results.csv"
degs_TOF_BINF_NR_w8vw0_own_csv <- "output/260519/rna_Tissue_NR_Inflamed_Week_8_vs_Baseline_deseq2_results.csv"
degs_TOF_BINF_RvNR_w0_own_csv <- "output/260519/rna_Tissue_Inflamed_Baseline_R_vs_NR_deseq2_results.csv"
degs_TOF_BINF_R_w8vw0_melonardanaz2025_csv <- "output/260615/rna_Tissue_R_Inflamed_Week_8_vs_Baseline_deseq2_results_MelonArdanaz2025.csv"
degs_TOF_BINF_NR_w8vw0_melonardanaz2025_csv <- "output/260615/rna_Tissue_NR_Inflamed_Week_8_vs_Baseline_deseq2_results_MelonArdanaz2025.csv"
degs_TOF_BINF_RvNR_w0_melonardanaz2025_csv <- "output/260615/rna_Tissue_Baseline_R_vs_NR_Inflamed_deseq2_results_MelonArdanaz2025.csv"
degs_IFX_BINF_R_w8vw0_arijs2009_csv <- "output/260519/Infliximab_R_Week_8_vs_Baseline_limma_results.csv"
degs_IFX_BINF_NR_w8vw0_arijs2009_csv <- "output/260519/Infliximab_NR_Week_8_vs_Baseline_limma_results.csv"
degs_IFX_BINF_RvNR_w0_arijs2009_csv <- "output/260519/Infliximab_R_vs_NR_Baseline_limma_results.csv"
degs_TOF_PBL_R_w8vw0_own_csv <- "output/260519/rna_PBL_R_Inflamed_Week_8_vs_Baseline_deseq2_results.csv"
degs_TOF_PBL_NR_w8vw0_own_csv <- "output/260519/rna_PBL_NR_Inflamed_Week_8_vs_Baseline_deseq2_results.csv"
degs_TOF_PBL_RvNR_w0_own_csv <- "output/260519/rna_PBL_Inflamed_Baseline_R_vs_NR_deseq2_results.csv"

olink_TOF_BINF_R_w8vw0_own_csv <- "output/260519/olink_Tissue_R_Inflamed_Week_8_vs_Baseline_limma_results.csv"
olink_TOF_BINF_NR_w8vw0_own_csv <- "output/260519/olink_Tissue_NR_Inflamed_Week_8_vs_Baseline_limma_results.csv"
olink_TOF_BINF_RvNR_w0_own_csv <- "output/260519/olink_Tissue_Inflamed_Baseline_R_vs_NR_limma_results.csv"
olink_TOF_SERUM_R_w8vw0_own_csv <- "output/260519/olink_serum_R_Inflamed_Week_8_vs_Baseline_limma_results.csv"
olink_TOF_SERUM_NR_w8vw0_own_csv <- "output/260519/olink_serum_NR_Inflamed_Week_8_vs_Baseline_limma_results.csv"
olink_TOF_SERUM_RvNR_w0_own_csv <- "output/260519/olink_serum_Inflamed_Baseline_R_vs_NR_limma_results.csv"

fgsea_TOF_BINF_R_w8vw0_own_csv <- "output/260519/rna_Tissue_R_Inflamed_Week_8_vs_Baseline_fgsea_results.csv"
fgsea_TOF_BINF_NR_w8vw0_own_csv <- "output/260519/rna_Tissue_NR_Inflamed_Week_8_vs_Baseline_fgsea_results.csv"
fgsea_TOF_BINF_RvNR_w0_own_csv <- "output/260519/rna_Tissue_Inflamed_Baseline_R_vs_NR_fgsea_results.csv"
fgsea_TOF_BINF_R_w8vw0_melonardanaz2025_csv <- "output/260615/rna_Tissue_R_Inflamed_Week_8_vs_Baseline_fgsea_results_MelonArdanaz2025.csv"
fgsea_TOF_BINF_NR_w8vw0_melonardanaz2025_csv <- "output/260615/rna_Tissue_NR_Inflamed_Week_8_vs_Baseline_fgsea_results_MelonArdanaz2025.csv"
fgsea_TOF_BINF_RvNR_w0_melonardanaz2025_csv <- "output/260615/rna_Tissue_Baseline_R_vs_NR_Inflamed_fgsea_results_MelonArdanaz2025.csv"
fgsea_IFX_BINF_R_w8vw0_arijs2009_csv <- "output/260519/Infliximab_R_Week_8_vs_Baseline_fgsea_results.csv"
fgsea_IFX_BINF_NR_w8vw0_arijs2009_csv <- "output/260519/Infliximab_NR_Week_8_vs_Baseline_fgsea_results.csv"
fgsea_IFX_BINF_RvNR_w0_arijs2009_csv <- "output/260519/Infliximab_R_vs_NR_Baseline_fgsea_results.csv"
fgsea_TOF_PBL_R_w8vw0_own_csv <- "output/260519/rna_PBL_R_Inflamed_Week_8_vs_Baseline_fgsea_results.csv"
fgsea_TOF_PBL_NR_w8vw0_own_csv <- "output/260519/rna_PBL_NR_Inflamed_Week_8_vs_Baseline_fgsea_results.csv"
fgsea_TOF_PBL_RvNR_w0_own_csv <- "output/260519/rna_PBL_Inflamed_Baseline_R_vs_NR_fgsea_results.csv"
```

```{r import, include = FALSE}
degs_TOF_BINF_R_w8vw0_own <- read.csv(degs_TOF_BINF_R_w8vw0_own_csv)
degs_TOF_BINF_NR_w8vw0_own <- read.csv(degs_TOF_BINF_NR_w8vw0_own_csv)
degs_TOF_BINF_RvNR_w0_own <- read.csv(degs_TOF_BINF_RvNR_w0_own_csv)
degs_TOF_BINF_R_w8vw0_melonardanaz2025 <- read.csv(degs_TOF_BINF_R_w8vw0_melonardanaz2025_csv)
degs_TOF_BINF_NR_w8vw0_melonardanaz2025 <- read.csv(degs_TOF_BINF_NR_w8vw0_melonardanaz2025_csv)
degs_TOF_BINF_RvNR_w0_melonardanaz2025 <- read.csv(degs_TOF_BINF_RvNR_w0_melonardanaz2025_csv)
degs_IFX_BINF_R_w8vw0_arijs2009 <- read.csv(degs_IFX_BINF_R_w8vw0_arijs2009_csv)
degs_IFX_BINF_NR_w8vw0_arijs2009 <- read.csv(degs_IFX_BINF_NR_w8vw0_arijs2009_csv)
degs_IFX_BINF_RvNR_w0_arijs2009 <- read.csv(degs_IFX_BINF_RvNR_w0_arijs2009_csv)
degs_TOF_PBL_R_w8vw0_own <- read.csv(degs_TOF_PBL_R_w8vw0_own_csv)
degs_TOF_PBL_NR_w8vw0_own <- read.csv(degs_TOF_PBL_NR_w8vw0_own_csv)
degs_TOF_PBL_RvNR_w0_own <- read.csv(degs_TOF_PBL_RvNR_w0_own_csv)

olink_TOF_BINF_R_w8vw0_own <- read.csv(olink_TOF_BINF_R_w8vw0_own_csv)
olink_TOF_BINF_NR_w8vw0_own <- read.csv(olink_TOF_BINF_NR_w8vw0_own_csv)
olink_TOF_BINF_RvNR_w0_own <- read.csv(olink_TOF_BINF_RvNR_w0_own_csv)
olink_TOF_SERUM_R_w8vw0_own <- read.csv(olink_TOF_SERUM_R_w8vw0_own_csv)
olink_TOF_SERUM_NR_w8vw0_own <- read.csv(olink_TOF_SERUM_NR_w8vw0_own_csv)
olink_TOF_SERUM_RvNR_w0_own <- read.csv(olink_TOF_SERUM_RvNR_w0_own_csv)

fgsea_TOF_BINF_R_w8vw0_own <- read.csv(fgsea_TOF_BINF_R_w8vw0_own_csv)
fgsea_TOF_BINF_NR_w8vw0_own <- read.csv(fgsea_TOF_BINF_NR_w8vw0_own_csv)
fgsea_TOF_BINF_RvNR_w0_own <- read.csv(fgsea_TOF_BINF_RvNR_w0_own_csv)
fgsea_TOF_BINF_R_w8vw0_melonardanaz2025 <- read.csv(fgsea_TOF_BINF_R_w8vw0_melonardanaz2025_csv)
fgsea_TOF_BINF_NR_w8vw0_melonardanaz2025 <- read.csv(fgsea_TOF_BINF_NR_w8vw0_melonardanaz2025_csv)
fgsea_TOF_BINF_RvNR_w0_melonardanaz2025 <- read.csv(fgsea_TOF_BINF_RvNR_w0_melonardanaz2025_csv)
fgsea_IFX_BINF_R_w8vw0_arijs2009 <- read.csv(fgsea_IFX_BINF_R_w8vw0_arijs2009_csv)
fgsea_IFX_BINF_NR_w8vw0_arijs2009 <- read.csv(fgsea_IFX_BINF_NR_w8vw0_arijs2009_csv)
fgsea_IFX_BINF_RvNR_w0_arijs2009 <- read.csv(fgsea_IFX_BINF_RvNR_w0_arijs2009_csv)
fgsea_TOF_PBL_R_w8vw0_own <- read.csv(fgsea_TOF_PBL_R_w8vw0_own_csv)
fgsea_TOF_PBL_NR_w8vw0_own <- read.csv(fgsea_TOF_PBL_NR_w8vw0_own_csv)
fgsea_TOF_PBL_RvNR_w0_own <- read.csv(fgsea_TOF_PBL_RvNR_w0_own_csv)
```

## Overlapping results

I want to identify which entities that are:
- Pretreatment response-associated in intestinal tissue, which change in a response-associated manner over time both. Preferably identify features that present are consistenly different at both gene and protein level.
- Pretreatment response-associated in intestinal tissue

```{r features}
degs_TOF_BINF_RvNR_w8vw0 <- degs_TOF_BINF_RvNR_w0_own %>%
  dplyr::select(gene_id, gene_symbol, log2FoldChange, stat, pvalue, padj) %>%
  dplyr::rename_at(vars(-gene_id, -gene_symbol), ~ paste0(., '_RvNR_w0')) %>%
  dplyr::inner_join(degs_TOF_BINF_R_w8vw0_own %>%
                      dplyr::select(gene_id, gene_symbol, log2FoldChange, stat, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene_id, -gene_symbol), ~ paste0(., '_R_w8vw0')), 
                    by = c("gene_id", "gene_symbol")) %>%
  dplyr::inner_join(degs_TOF_BINF_NR_w8vw0_own %>%
                      dplyr::select(gene_id, gene_symbol, log2FoldChange, stat, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene_id, -gene_symbol), ~ paste0(., '_NR_w8vw0')), 
                    by = c("gene_id", "gene_symbol")) %>%
  dplyr::mutate(significance = case_when(
    padj_RvNR_w0 < 0.05 & padj_R_w8vw0 < 0.05 ~ "Pretreatment, response-associated",
    padj_RvNR_w0 < 0.05 & padj_NR_w8vw0 < 0.05 ~ "Pretreatment, non-response-associated",
    padj_RvNR_w0 > 0.05 & padj_R_w8vw0 < 0.05 ~ "Response-associated",
    padj_RvNR_w0 > 0.05 & padj_NR_w8vw0 < 0.05 ~ "Non-response-associated",
    .default = "NS",
  ))

degs_TOF_BINF_RvNR_w8vw0_melonardanaz2025 <- degs_TOF_BINF_RvNR_w0_melonardanaz2025 %>%
  dplyr::select(gene, log2FoldChange, stat, pvalue, padj) %>%
  dplyr::rename_at(vars(-gene), ~ paste0(., '_RvNR_w0')) %>%
  dplyr::inner_join(degs_TOF_BINF_R_w8vw0_melonardanaz2025 %>%
                      dplyr::select(gene, log2FoldChange, stat, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene), ~ paste0(., '_R_w8vw0')), 
                    by = c("gene")) %>%
  dplyr::inner_join(degs_TOF_BINF_NR_w8vw0_melonardanaz2025 %>%
                      dplyr::select(gene, log2FoldChange, stat, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene), ~ paste0(., '_NR_w8vw0')), 
                    by = c("gene")) %>%
  dplyr::mutate(significance = case_when(
    padj_RvNR_w0 < 0.05 & padj_R_w8vw0 < 0.05 ~ "Pretreatment, response-associated",
    padj_RvNR_w0 < 0.05 & padj_NR_w8vw0 < 0.05 ~ "Pretreatment, non-response-associated",
    padj_RvNR_w0 > 0.05 & padj_R_w8vw0 < 0.05 ~ "Response-associated",
    padj_RvNR_w0 > 0.05 & padj_NR_w8vw0 < 0.05 ~ "Non-response-associated",
    .default = "NS",
  ))

olink_TOF_BINF_RvNR_w8vw0 <-  olink_TOF_BINF_RvNR_w0_own %>%
  dplyr::select(gene_symbol, log2FoldChange, t, pvalue, padj) %>%
  dplyr::rename_at(vars(-gene_symbol), ~ paste0(., '_RvNR_w0')) %>%
  dplyr::inner_join(olink_TOF_BINF_R_w8vw0_own %>%
                      dplyr::select(gene_symbol, log2FoldChange, t, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene_symbol), ~ paste0(., '_R_w8vw0')), 
                    by = c("gene_symbol")) %>%
  dplyr::inner_join(olink_TOF_BINF_NR_w8vw0_own %>%
                      dplyr::select(gene_symbol, log2FoldChange, t, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene_symbol), ~ paste0(., '_NR_w8vw0')), 
                    by = c("gene_symbol")) %>%
  dplyr::mutate(significance = case_when(
    padj_RvNR_w0 < 0.05 & padj_R_w8vw0 < 0.05 ~ "Pretreatment, response-associated",
    padj_RvNR_w0 < 0.05 & padj_NR_w8vw0 < 0.05 ~ "Pretreatment, non-response-associated",
    padj_RvNR_w0 > 0.05 & padj_R_w8vw0 < 0.05 ~ "Response-associated",
    padj_RvNR_w0 > 0.05 & padj_NR_w8vw0 < 0.05 ~ "Non-response-associated",
    .default = "NS",
  ))

degs_TOF_PBL_RvNR_w8vw0 <- degs_TOF_PBL_RvNR_w0_own %>%
  dplyr::select(gene_id, gene_symbol, log2FoldChange, stat, pvalue, padj) %>%
  dplyr::rename_at(vars(-gene_id, -gene_symbol), ~ paste0(., '_RvNR_w0')) %>%
  dplyr::inner_join(degs_TOF_PBL_R_w8vw0_own %>%
                      dplyr::select(gene_id, gene_symbol, log2FoldChange, stat, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene_id, -gene_symbol), ~ paste0(., '_R_w8vw0')), 
                    by = c("gene_id", "gene_symbol")) %>%
  dplyr::inner_join(degs_TOF_PBL_NR_w8vw0_own %>%
                      dplyr::select(gene_id, gene_symbol, log2FoldChange, stat, pvalue, padj) %>%
                      dplyr::rename_at(vars(-gene_id, -gene_symbol), ~ paste0(., '_NR_w8vw0')), 
                    by = c("gene_id", "gene_symbol")) %>%
  dplyr::mutate(significance = case_when(
    padj_RvNR_w0 < 0.05 & padj_R_w8vw0 < 0.05 ~ "Pretreatment, response-associated",
    padj_RvNR_w0 < 0.05 & padj_NR_w8vw0 < 0.05 ~ "Pretreatment, non-response-associated",
    padj_RvNR_w0 > 0.05 & padj_R_w8vw0 < 0.05 ~ "Response-associated",
    padj_RvNR_w0 > 0.05 & padj_NR_w8vw0 < 0.05 ~ "Non-response-associated",
    .default = "NS",
  ))
```

```{r pathways wide}
fgsea_TOF_BINF_RvNR_w8vw0_own_wide <- fgsea_TOF_BINF_R_w8vw0_own %>%
  dplyr::mutate(significance = case_when(
    padj<0.05 & NES<0 ~ "SS Down",
    padj<0.05 & NES>0 ~ "SS Up",
    pval<0.05 & NES<0 ~ "S Down",
    pval<0.05 & NES>0 ~ "S Up",
    pval>0.05 & NES<0 ~ "NS Down",
    pval>0.05 & NES>0 ~ "NS Up",
  )) %>%
  dplyr::select(pathway, pval, padj, NES, significance) %>%
  dplyr::rename_at(vars(-pathway), ~ paste0(., '_TOF_BINF_R_w8vw0_own')) %>%
  dplyr::inner_join(fgsea_TOF_BINF_NR_w8vw0_own %>%
                      dplyr::mutate(significance = case_when(
                        padj<0.05 & NES<0 ~ "SS Down",
                        padj<0.05 & NES>0 ~ "SS Up",
                        pval<0.05 & NES<0 ~ "S Down",
                        pval<0.05 & NES>0 ~ "S Up",
                        pval>0.05 & NES<0 ~ "NS Down",
                        pval>0.05 & NES>0 ~ "NS Up",
                      )) %>%
                      dplyr::select(pathway, pval, padj, NES, significance) %>%
                      dplyr::rename_at(vars(-pathway), ~ paste0(., '_TOF_BINF_NR_w8vw0_own')),
                    by = "pathway")

fgsea_TOF_BINF_RvNR_w8vw0_melonardanaz2025_wide <- fgsea_TOF_BINF_R_w8vw0_melonardanaz2025 %>%
  dplyr::mutate(significance = case_when(
    padj<0.05 & NES<0 ~ "SS Down",
    padj<0.05 & NES>0 ~ "SS Up",
    pval<0.05 & NES<0 ~ "S Down",
    pval<0.05 & NES>0 ~ "S Up",
    pval>0.05 & NES<0 ~ "NS Down",
    pval>0.05 & NES>0 ~ "NS Up",
  )) %>%
  dplyr::select(pathway, pval, padj, NES, significance) %>%
  dplyr::rename_at(vars(-pathway), ~ paste0(., '_TOF_BINF_R_w8vw0_am2025')) %>%
  dplyr::inner_join(fgsea_TOF_BINF_NR_w8vw0_melonardanaz2025 %>%
                      dplyr::mutate(significance = case_when(
                        padj<0.05 & NES<0 ~ "SS Down",
                        padj<0.05 & NES>0 ~ "SS Up",
                        pval<0.05 & NES<0 ~ "S Down",
                        pval<0.05 & NES>0 ~ "S Up",
                        pval>0.05 & NES<0 ~ "NS Down",
                        pval>0.05 & NES>0 ~ "NS Up",)) %>%
                      dplyr::select(pathway, pval, padj, NES, significance) %>%
                      dplyr::rename_at(vars(-pathway), ~ paste0(., '_TOF_BINF_NR_w8vw0_am2025')),
                    by = "pathway")

fgsea_IFX_BINF_RvNR_w8vw0_arijs2009_wide <- fgsea_IFX_BINF_R_w8vw0_arijs2009 %>%
  dplyr::mutate(significance = case_when(
    padj<0.05 & NES<0 ~ "SS Down",
    padj<0.05 & NES>0 ~ "SS Up",
    pval<0.05 & NES<0 ~ "S Down",
    pval<0.05 & NES>0 ~ "S Up",
    pval>0.05 & NES<0 ~ "NS Down",
    pval>0.05 & NES>0 ~ "NS Up",
  )) %>%
  dplyr::select(pathway, pval, padj, NES, significance) %>%
  dplyr::rename_at(vars(-pathway), ~ paste0(., '_IFX_BINF_R_w8vw0_arijs2009')) %>%
  dplyr::inner_join(fgsea_IFX_BINF_NR_w8vw0_arijs2009 %>%
                      dplyr::mutate(significance = case_when(
                        padj<0.05 & NES<0 ~ "SS Down",
                        padj<0.05 & NES>0 ~ "SS Up",
                        pval<0.05 & NES<0 ~ "S Down",
                        pval<0.05 & NES>0 ~ "S Up",
                        pval>0.05 & NES<0 ~ "NS Down",
                        pval>0.05 & NES>0 ~ "NS Up",
                      )) %>%
                      dplyr::select(pathway, pval, padj, NES, significance) %>%
                      dplyr::rename_at(vars(-pathway), ~ paste0(., '_IFX_BINF_NR_w8vw0_arijs2009')),
                    by = "pathway")

fgsea_TOF_PBL_RvNR_w8vw0_own <- fgsea_TOF_PBL_R_w8vw0_own %>%
  dplyr::mutate(significance = case_when(
    padj<0.05 & NES<0 ~ "SS Down",
    padj<0.05 & NES>0 ~ "SS Up",
    pval<0.05 & NES<0 ~ "S Down",
    pval<0.05 & NES>0 ~ "S Up",
    pval>0.05 & NES<0 ~ "NS Down",
    pval>0.05 & NES>0 ~ "NS Up",
  )) %>%
  dplyr::select(pathway, pval, padj, NES, significance) %>%
  dplyr::rename_at(vars(-pathway), ~ paste0(., '_TOF_PBL_R_w8vw0_own')) %>%
  dplyr::inner_join(fgsea_TOF_PBL_NR_w8vw0_own %>%
                      dplyr::mutate(significance = case_when(
                        padj<0.05 & NES<0 ~ "SS Down",
                        padj<0.05 & NES>0 ~ "SS Up",
                        pval<0.05 & NES<0 ~ "S Down",
                        pval<0.05 & NES>0 ~ "S Up",
                        pval>0.05 & NES<0 ~ "NS Down",
                        pval>0.05 & NES>0 ~ "NS Up",)) %>%
                      dplyr::select(pathway, pval, padj, NES, significance) %>%
                      dplyr::rename_at(vars(-pathway), ~ paste0(., '_TOF_PBL_NR_w8vw0_own')),
                    by = "pathway")

fgsea_total_wide <- fgsea_TOF_BINF_RvNR_w8vw0_own_wide %>%
  dplyr::inner_join(fgsea_TOF_BINF_RvNR_w8vw0_melonardanaz2025_wide,
                    by = "pathway") %>%
  dplyr::inner_join(fgsea_IFX_BINF_RvNR_w8vw0_arijs2009_wide,
                    by = "pathway") %>%
  dplyr::inner_join(fgsea_TOF_PBL_RvNR_w8vw0_own,
                    by = "pathway")

fgsea_total_wide_sig <- fgsea_total_wide %>%
  dplyr::filter(significance_TOF_BINF_R_w8vw0_own %in% c("SS Up", "SS Down"),
                significance_TOF_BINF_R_w8vw0_am2025 == significance_TOF_BINF_R_w8vw0_own) %>%
  dplyr::mutate(significance = case_when(
    (significance_TOF_BINF_R_w8vw0_own=="SS Up" & 
      significance_TOF_BINF_NR_w8vw0_own%in%c("NS Up", "NS Down", "SS Down", "S Down") & 
      significance_IFX_BINF_R_w8vw0_arijs2009%in%c("S Down", "SS Down")) ~ "TOF R up; IFX R down",
    (significance_TOF_BINF_R_w8vw0_own=="SS Down" & 
      significance_TOF_BINF_NR_w8vw0_own%in%c("NS Up", "NS Down", "SS Up", "S Up") & 
      significance_IFX_BINF_R_w8vw0_arijs2009 %in%c("S Up", "SS Up")) ~ "TOF R down; IFX R up",
    (significance_TOF_BINF_R_w8vw0_own=="SS Up" & 
      significance_TOF_BINF_NR_w8vw0_own%in%c("NS Up", "NS Down", "SS Down", "S Down") & 
      significance_IFX_BINF_R_w8vw0_arijs2009%in%c("S Up", "SS Up")) ~ "TOF R up; IFX R up",
    (significance_TOF_BINF_R_w8vw0_own=="SS Down" & 
      significance_TOF_BINF_NR_w8vw0_own%in%c("NS Up", "NS Down", "SS Up", "S Up") & 
      significance_IFX_BINF_R_w8vw0_arijs2009%in%c("S Down", "SS Down")) ~ "TOF R down; IFX R down",
    (significance_TOF_BINF_R_w8vw0_own=="SS Up" & 
       significance_TOF_BINF_NR_w8vw0_own%in%c("NS Up", "NS Down", "SS Down", "S Down") & 
       significance_IFX_BINF_R_w8vw0_arijs2009%in%c("NS Up", "NS Down")) ~ "TOF R up",
    (significance_TOF_BINF_R_w8vw0_own=="SS Down" & 
       significance_TOF_BINF_NR_w8vw0_own%in%c("NS Up", "NS Down", "SS Up", "S Up") & 
       significance_IFX_BINF_R_w8vw0_arijs2009%in%c("NS Up", "NS Down")) ~ "TOF R down",
    .default="NS")) %>%
  dplyr::select(pathway, starts_with("significance"))

fgsea_total_wide_sig %>%
  dplyr::filter(significance %in% c("TOF R down", "TOF R up"))

fgsea_total_wide_interesting <- fgsea_total_wide %>%
  dplyr::filter_out(grepl("^NS", significance_TOF_BINF_R_w8vw0_own) & 
                      grepl("^NS", significance_TOF_BINF_NR_w8vw0_own) & 
                      grepl("^NS", significance_TOF_BINF_R_w8vw0_am2025) & 
                      grepl("^NS", significance_TOF_BINF_NR_w8vw0_am2025) &
                      grepl("^NS", significance_IFX_BINF_R_w8vw0_arijs2009) &
                      grepl("^NS", significance_IFX_BINF_NR_w8vw0_arijs2009)) %>%
  dplyr::mutate(significance = case_when(
    significance_TOF_BINF_R_w8vw0_own == "Up" & 
      significance_TOF_BINF_R_w8vw0_am2025 == "Up" & 
      significance_TOF_BINF_NR_w8vw0_own %in% c("NS Up", "NS Down", "Down") & 
      significance_IFX_BINF_R_w8vw0_arijs2009 %in% c("NS Up", "NS Down", "Down") ~ "Tofacitinib-specific, concordant response-associated upregulation",
    significance_TOF_BINF_R_w8vw0_own == "Down" & 
      significance_TOF_BINF_R_w8vw0_am2025 == "Down" & 
      significance_TOF_BINF_NR_w8vw0_own %in% c("NS Up", "NS Down", "Up") & 
      significance_IFX_BINF_R_w8vw0_arijs2009 %in% c("NS Up", "NS Down", "Up") ~ "Tofacitinib-specific, concordant response-associated downregulation"
  ),
  )

fgsea_total_wide_conctofspec <- fgsea_total_wide_interesting %>%
  dplyr::filter(significance != "NS") %>%
  dplyr::arrange(significance)

```

# Tofacitinib response-associated differences, validated by Melon-Ardanaz, but opposite in infliximab responders

Which pathways are down in tofactinib responders and non-responders.

```{r}
fgsea_TOF_BINF_RvNR_w8vw0_own_wide %>% 
  dplyr::filter(significance_TOF_BINF_R_w8vw0_own == "Down" & significance_TOF_BINF_NR_w8vw0_own == "Down")
```

```{r}
fgsea_TOF_BINF_RvNR_w8vw0_own_wide %>% 
  dplyr::filter(significance_TOF_BINF_R_w8vw0_own == "Down" & significance_TOF_BINF_NR_w8vw0_own != "Down")
```

```{r JAK pathways}
fgsea_TOF_BINF_RvNR_w8vw0_own_wide %>%
  dplyr::filter(grepl("JAK", pathway))
```

```{r JAK pathways}
fgsea_TOF_BINF_RvNR_w8vw0_own_wide %>%
  dplyr::filter(grepl("INTERFERON", pathway))
```

Which pathways are discordant between tofactinib and infliximab responders.

```{r}
fgsea_total_wide_conctofspec_tofifxdisc <- fgsea_total_wide_conctofspec %>%
  dplyr::mutate(direction_IFX_BINF_R_w8vw0_arijs2009 = gsub("NS ", "", significance_IFX_BINF_R_w8vw0_arijs2009)) %>%
  dplyr::filter(significance_TOF_BINF_R_w8vw0_own != direction_IFX_BINF_R_w8vw0_arijs2009)
```


```{r pathways long}
fgsea_TOF_BINF_RvNR_w8vw0_own_long <- fgsea_TOF_BINF_R_w8vw0_own %>%
  dplyr::select(pathway, pval, padj, NES) %>%
  dplyr::mutate(Response = "R") %>%
  dplyr::rows_append(fgsea_TOF_BINF_NR_w8vw0_own %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "NR")) %>%
  dplyr::rows_append(fgsea_TOF_BINF_RvNR_w0_own %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "RvNR")) %>%
  dplyr::mutate(Treatment = "Tofacitinib",
                Tissue = "BINF",
                Source = "Own")

fgsea_TOF_BINF_RvNR_w8vw0_melonardanaz2025_long <- fgsea_TOF_BINF_R_w8vw0_melonardanaz2025 %>%
  dplyr::select(pathway, pval, padj, NES) %>%
  dplyr::mutate(Response = "R") %>%
  dplyr::rows_append(fgsea_TOF_BINF_NR_w8vw0_melonardanaz2025 %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "NR")) %>%
  dplyr::rows_append(fgsea_TOF_BINF_RvNR_w0_melonardanaz2025 %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "RvNR")) %>%
  dplyr::mutate(Treatment = "Tofacitinib",
                Tissue = "BINF",
                Source = "Melon-Ardanaz 2025")

fgsea_IFX_BINF_RvNR_w8vw0_arijs2009_long <- fgsea_IFX_BINF_R_w8vw0_arijs2009 %>%
  dplyr::select(pathway, pval, padj, NES) %>%
  dplyr::mutate(Response = "R") %>%
  dplyr::rows_append(fgsea_IFX_BINF_NR_w8vw0_arijs2009 %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "NR")) %>%
  dplyr::rows_append(fgsea_IFX_BINF_RvNR_w0_arijs2009 %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "RvNR")) %>%
  dplyr::mutate(Treatment = "Infliximab",
                Tissue = "BINF",
                Source = "Arijs 2009")

fgsea_TOF_PBL_RvNR_w8vw0_own_long <- fgsea_TOF_PBL_R_w8vw0_own %>%
  dplyr::select(pathway, pval, padj, NES) %>%
  dplyr::mutate(Response = "R") %>%
  dplyr::rows_append(fgsea_TOF_PBL_NR_w8vw0_own %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "NR")) %>%
  dplyr::rows_append(fgsea_TOF_PBL_RvNR_w0_own %>%
                       dplyr::select(pathway, pval, padj, NES) %>%
                       dplyr::mutate(Response = "RvNR")) %>%
  dplyr::mutate(Treatment = "Tofacitinib",
                Tissue = "PBL",
                Source = "Own")

fgsea_total_long <- fgsea_TOF_BINF_RvNR_w8vw0_own_long %>%
  dplyr::rows_append(fgsea_TOF_BINF_RvNR_w8vw0_melonardanaz2025_long) %>%
  dplyr::rows_append(fgsea_IFX_BINF_RvNR_w8vw0_arijs2009_long) %>%
  dplyr::rows_append(fgsea_TOF_PBL_RvNR_w8vw0_own_long)
```


```{r, fig.height = 12, fig.width = 25}
arrowplot_fgsea_total_long_ggplotobj <- fgsea_total_long %>%
  dplyr::filter(pathway %in% (fgsea_total_wide_sig %>%
                                dplyr::filter(significance %in% c("TOF R down", "TOF R up")) %>%
                                dplyr::pull(pathway))) %>%
  dplyr::mutate(Direction = ifelse(NES<0, "Down", "Up"),
                Significance = ifelse(padj<0.05, "Significant", "NS")) %>%
  ggplot(aes(x = pathway, y = Source)) +
  geom_point(aes(
    shape = Direction,
    fill = NES,
    size = -log10(padj),
    alpha = Significance)) +
  facet_grid(Response+Tissue~., scales = "free", space = "free") +
  labs(title = "NES",
       subtitle = "Week 8 relative to Week 0") +
  scale_alpha_discrete(range = c(0.25, 1)) +
  scale_shape_manual(values = c("Up" = 24, "Down" = 25)) +
  scale_fill_gradient2(high = "coral3", low = "deepskyblue3", mid = "white") +
  # scale_fill_gradient2(high = "#1c8859", low = "#a32a2f", mid = "white") +
  theme_bw() +
  guides(size = guide_legend(override.aes = list(shape=17))) +
  theme(axis.title.y =  element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, face = "bold"),
        axis.text.y = element_text(face = "bold"),
        # plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.title = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))

print(arrowplot_fgsea_total_long_ggplotobj)

ggsave("arrowplot_fgsea_conctofspec_long.pdf", width = 25, height = 11)
```

```{r}
fgsea_TOF_BINF_RvNR_w8vw0_ownvmelonardanaz2025 %>%
  dplyr::filter(significance != "NS") %>%
  dplyr::select(pathway)
```

## RvNR: w8vw0: Own

```{r fgsea rvnr w8vw0 own}
fgsea_TOF_BINF_R_w8vw0_own_sig <- fgsea_TOF_BINF_R_w8vw0_own %>%
  dplyr::filter(padj<0.05)

fgsea_TOF_BINF_NR_w8vw0_own_Rsigfilt_sig <- fgsea_TOF_BINF_NR_w8vw0_own %>%
  dplyr::filter(pathway %in% fgsea_TOF_BINF_R_w8vw0_own_sig$pathway) %>%
  dplyr::arrange(pval) %>%
  dplyr::filter(padj<0.05)
```

```{r}
fgsea_TOF_BINF_RvNR_w8vw0_own_wide_treatment <- fgsea_TOF_BINF_RvNR_w8vw0_own_wide %>%
  dplyr::filter(significance_TOF_BINF_R_w8vw0_own == significance_TOF_BINF_NR_w8vw0_own,
                significance_TOF_BINF_R_w8vw0_own != "NS") %>%
  dplyr::arrange(significance_TOF_BINF_R_w8vw0_own)


fgsea_TOF_BINF_RvNR_w8vw0_own_wide_treatment %>%
  dplyr::filter(pathway %in% c("HALLMARK_COMPLEMENT", "HALLMARK_INFLAMMATORY_RESPONSE", "HALLMARK_INTERFERON_ALPHA_RESPONSE", "HALLMARK_INTERFERON_GAMMA_RESPONSE"))

fgsea_TOF_BINF_RvNR_w8vw0_own_wide_treatment %>%
  dplyr::filter(significance_TOF_BINF_R_w8vw0_own == "Up")
```


## RvNR: w8vw0: OwnvMelonArdanaz2025

### Genes

```{r degs rvnr w8vw0 consistent}
degs_BINF_R_w8vw0_consistent <- degs_BINF_R_w8vw0 %>%
  dplyr::inner_join(degs_BINF_R_w8vw0_melonardanaz2025, by = c("gene_symbol" = "gene"), suffix = c("_own", "_am2025")) %>%
  dplyr::mutate(Response = "R")

degs_BINF_NR_w8vw0_consistent <- degs_BINF_NR_w8vw0 %>%
  dplyr::inner_join(degs_BINF_NR_w8vw0_melonardanaz2025, by = c("gene_symbol" = "gene"), suffix = c("_own", "_am2025")) %>%
  dplyr::mutate(Response = "NR")

degs_BINF_RvNR_w8vw0_consistent <- degs_BINF_R_w8vw0_consistent %>%
  dplyr::rows_append(degs_BINF_NR_w8vw0_consistent) %>%
  dplyr::mutate(Response = factor(Response, levels = c("R", "NR")))
```

```{r degs r w8vw0 consistent sig}
degs_BINF_R_w8vw0_consistent_sig <- degs_BINF_R_w8vw0_consistent %>%
  dplyr::filter(pvalue_own < 0.05 & pvalue_am2025 < 0.05) %>%
  dplyr::mutate(significance = case_when(
    stat_own < 0 & stat_am2025 < 0 ~ "Down",
    stat_own > 0 & stat_am2025 > 0 ~ "Up",
    .default = "Inconsistent"
  ))

degs_BINF_NR_w8vw0_consistent_sig <- degs_BINF_NR_w8vw0_consistent %>%
  dplyr::filter(pvalue_own < 0.05 & pvalue_am2025 < 0.05) %>%
  dplyr::mutate(significance = case_when(
    stat_own < 0 & stat_am2025 < 0 ~ "Down",
    stat_own > 0 & stat_am2025 > 0 ~ "Up",
    .default = "Inconsistent"
  ))
```

```{r degs rvnr w8vw0 consistent plot, fig.width = 9, fig.height = 5}
degs_BINF_RvNR_w8vw0_consistent %>%
  ggplot(aes(x = log2FoldChange_own, y = log2FoldChange_am2025)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point_rast() +
  geom_smooth(method = "lm") +
  labs(title = "Comparison with Ardanaz Melon 2025\nDOI: 10.1093/ecco-jcc/jjaf076",
       subtitle = "log2 (fold change)",
       x = "Own\n(bulk RNAseq)",
       y = "Ardanaz Melon 2025\n(pseudobulk scRNAseq)") +
  facet_grid(.~Response) +
  theme_bw() +
  xlim(c(-10, 10)) +
  ylim(c(-10, 10)) +
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_blank())
```

Maybe good idea to highlight genes that were found to be TOF-specific (i.e. not found in IFX)

### Pathways

```{r fgsea rvnr w8vw0 consistent}
fgsea_BINF_R_w8vw0_consistent <- fgsea_TOF_BINF_R_w8vw0_consistent %>%
  dplyr::inner_join(fgsea_BINF_R_w8vw0_melonardanaz2025, by = c("pathway"), suffix = c("_own", "_am2025")) %>%
  dplyr::mutate(Response = "R")

fgsea_BINF_NR_w8vw0_consistent <- fgsea_BINF_NR_w8vw0 %>%
  dplyr::inner_join(fgsea_BINF_NR_w8vw0_melonardanaz2025, by = c("pathway"), suffix = c("_own", "_am2025")) %>%
  dplyr::mutate(Response = "NR")

fgsea_BINF_RvNR_w8vw0_consistent <- fgsea_BINF_R_w8vw0_consistent %>%
  dplyr::rows_append(fgsea_BINF_NR_w8vw0_consistent) %>%
  dplyr::mutate(Response = factor(Response, levels = c("R", "NR")))
```

```{r fgsea r w8vw0 consistent sig}
fgsea_BINF_R_w8vw0_consistent_sig <- fgsea_BINF_R_w8vw0_consistent %>%
  dplyr::filter(pval_own < 0.05 & pval_am2025 < 0.05) %>%
  dplyr::mutate(significance = case_when(
    NES_own < 0 & NES_am2025 < 0 ~ "Down",
    NES_own > 0 & NES_am2025 > 0 ~ "Up",
    .default = "Inconsistent"
  ))

fgsea_BINF_NR_w8vw0_consistent_sig <- fgsea_BINF_NR_w8vw0_consistent %>%
  dplyr::filter(pval_own < 0.05 & pval_am2025 < 0.05) %>%
  dplyr::mutate(significance = case_when(
    NES_own < 0 & NES_am2025 < 0 ~ "Down",
    NES_own > 0 & NES_am2025 > 0 ~ "Up",
    .default = "Inconsistent"
  ))
```

```{r fgsea rvnr w8vw0 consistent plot, fig.width = 9, fig.height = 5}
fgsea_BINF_RvNR_w8vw0_consistent %>%
  ggplot(aes(x = NES_own, y = NES_am2025)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point_rast() +
  geom_smooth(method = "lm") +
  labs(title = "Comparison with Ardanaz Melon 2025\nDOI: 10.1093/ecco-jcc/jjaf076",
       subtitle = "NES",
       x = "Own\n(bulk RNAseq)",
       y = "Ardanaz Melon 2025\n(pseudobulk scRNAseq)") +
  facet_grid(.~Response) +
  theme_bw() +
  xlim(c(-3, 3)) +
  ylim(c(-3, 3)) +
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_blank())
```

Maybe good idea to highlight genes that were found to be TOF-specific (i.e. not found in IFX)

